High-order cognition is supported by complex but compressible brain activity patterns

### 2023-02-23 paper/main.tex
working on intro

To the extent that brain activity patterns contain rich task-relevant
information, we should be able to use the activity patterns to accurately
differentiate between different aspects of the task~\citep[e.g., using pattern
classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
correspondence between classification accuracy and the information content of a
signal~\citep{Alva02}. To the extent that brain activity patterns are
compressible, we should be able to generate simplified (e.g., lower
dimensional) representations of the data while still preserving the relevant or
important aspects of the original signal. In general, information content and
compressibility are related but are partially dissociable
(Fig.~\ref{fig:information-compression}).

@@

% Specifically, if a given signal
% (e.g., a representation of brain activity patterns) contains more information
% about ongoing cognitive processes, then the peak decoding accuracy should be
% high. And if the signal is compressible, a low-dimensional embedding of the
% signal will be similarly informative to the original signal. A signal can be
% informative but not compressible, informative but

% %   - e.g., you could have something that is highly compressible (just a few components are sufficient to achieve some threshold information level) but
% %     *also* very rich-- e.g., the maximum number of informative components is large
% % compression and task demands -- MackEtal20, ZhouEtal19
% % superEEG -- prediction of held-out data; implies redundancy
% % timecorr -- higher order network patterns during higher order cognition implies more complexity
% % cognitive richness and scrambling-- various Hasson lab papers

>>>>

To the extent that brain activity patterns contain rich task-relevant
information, we should be able to use the activity patterns to accurately
differentiate between different aspects of the task~\citep[e.g., using pattern
classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
correspondence between classification accuracy and the information content of a
signal~\citep{Alva02}. To the extent that brain activity patterns are
compressible, we should be able to generate simplified (e.g., lower
dimensional) representations of the data while still preserving the relevant or
important aspects of the original signal. In general, information content and
compressibility are related but are partially dissociable
(Fig.~\ref{fig:information-compression}). If a given signal (e.g., a
representation of brain activity patterns) contains more information about
ongoing cognitive processes, then the peak decoding accuracy should be high.
And if the signal is compressible, a low-dimensional embedding of the signal
will be similarly informative to the original signal
(Fig.~\ref{fig:information-compression}D).

@@

Several recent studies suggest that the complexity of brain activity is
task-dependent, whereby simpler tasks with lower cognitive demands are
reflected by simpler and more compressible brain activity patterns, and more
complex tasks with higher cognitive demands are reflected by more complex and
less compressible brain activity patterns~\citep{MackEtal20, OwenEtal21}. These
findings complement other work suggesting that functional connectivity
(correlation) patterns are task-dependent~\citep{FinnEtal17, OwenEtal20,
ColeEtal14}, although see~\cite{GratEtal18}. Higher-order cognitive processing
of a common stimulus also appears to drive more stereotyped task-related
activity and functional connectivity across individuals~\citep{HassEtal08,
LernEtal11, SimoChan20, SimoEtal16}.

% % compression and task demands -- MackEtal20, ZhouEtal19
% % timecorr -- higher order network patterns during higher order cognition implies more complexity
% % cognitive richness and scrambling-- various Hasson lab papers
